\author{Samuel Gido}
\title{Week 1 Discrete}
\date{Spring 2024}

\documentclass[12pt, letterpaper]{article}

\usepackage{fullpage}
\usepackage{setspace}
\onehalfspacing

\begin{document} 
\maketitle

\section*{Language of Sets}

Let \(S\) and \(T\) be sets, these are basic set operations.
\begin{center}
  \( \vert S \vert\) = \#\( S\) = \# of elements in S \\
  \(S \cup T\) = \(S\) union \(T\) = \{x: x \(\in\) S or x \(\in\) T\} \\
  \(S \cap T\) = \(S\) intersection \(T\) = \{x: x \(\in\) S and x \(\in\) T\} \\
  \(\emptyset \) = empty set = \{ \} \\
  \(S \backslash T\) = \(S\) minus \(T\) = \{x: x \(\in\) S and x \(\notin\) T\} \\
  \(S \sqcup T\) = \(S\) symmetric difference \(T\) = \{x: x \(\in\) S or x \(\in\) T but not both\} (similar to xor) \\
  \(S \subset T\) = \(S\) is a subset of \(T\) = \{x: x \(\in\) S implies x \(\in\) T\} \\
  \(\bar{S}\) = \(S^{\subset}\) = complement of \(S\) = \{x: x \(\notin\) S\} \\
\end{center}

\subsection*{Partitioning a Set}
\indent Let \(S\) be a set. Partitioning \(S\) would create a collection \(S_1, S_2,...,S_m\) such that each subset of \(S\) is exactly one element of \(S\). Each subset is called a \textit{part}.

\begin{center}
  \(S = S_1 \cup S_2 \cup ... \cup S_m \) \\
  \(S_i \cap S_j = \emptyset, \,\,\, (i \neq j)\)
\end{center}

\pagebreak
\section*{Addition Principle}
\begin{center}
  \(\vert S \vert = \vert S_1 \vert + \vert S_2 \vert + ... + \vert S_m \vert\) \\
\end{center}

From the principle that a whole is equal to the sum of its parts. This only applies if the whole set is correctly partitioned, or the parts don't overlap. \\

\textbf{Example:} A bag has 5 red marbles, 3 blue marbles, and 2 green marbles. How many ways can you pick a marble from the bag? \\
\begin{center}
  \(\vert S \vert = \vert S_1 \vert + \vert S_2 \vert + \vert S_3 \vert = 5 + 3 + 2 = 10\) ways to pick a marble from the bag\\
\end{center}

\section*{Multiplication Principle}
\begin{center}
  Let \(S\) = \{(\(a, b\)) : \(\vert a \vert = p, \vert b \vert = q\)\} \\
  Then \(\vert S \vert = pq\)
\end{center}
This principle is a consequence of the addition principle. Let \(a_1, a_2,...,a_p\) be the partition of \(a\). Next, partition \(S\) into parts \(S_1, S_2,..., S_p\) where \(S_i\) is the set of ordered pairs in \(S\) with \(a_i\) as the first element (\(i = 1, 2,..., p\)). For each element in \(a\), there are \(q\) possible options to make a pair with that element in a. \\

\textbf{Example:} If there are 6 men, 4 women, 3 boys, and 2 girls, how many ways can you get one of each? \\
\begin{center}
  \(\vert S \vert = \vert S_1 \vert \cdot \vert S_2 \vert \cdot \vert S_3 \vert \cdot \vert S_4 \vert = 6 \cdot 4 \cdot 3 \cdot 2 = 144\) ways to get one of each.
\end{center}

\textbf{Example:} \(N = 2^9 \cdot 5^7 \cdot 7^3\), what is the number of positive numbers that are factors of \(N\)?

\begin{center}
  Consider \(m\) such that \(m \vert N\), \\
  For any factor, \(m = 2^a \cdot 5^b \cdot 7^c \), \\
  Where \(
    a = \{ 0 \dots 9 \}, \,b =\{0 \dots 7\}, \,c = \{0 \dots 3\}
  \) \\
  This means there are 10 options for \(a\), 8 options for \(b\), and 4 options for \(c\). \\
  Therefore, \(10 \cdot 8 \cdot 4 = 320\) positive integer factors of \(N\) exist.
\end{center}

\pagebreak
\section*{Subtraction Principle}
\begin{center}
  Let \(A\) be a set and let \(U\) be a set larger than and containing \(A\).
  \(\bar{A} = U / A = \{x \in U : x \notin A\}\) = the complement of \(A\) in \(U\). \\
  Therefore,\\
   \(\vert A \vert = \vert U \vert - \vert \bar{A} \vert\).
\end{center}
This principle only makes sense when working with objects in \(U\) and \(\bar{A}\) is easier than working with objects in \(A\). \\ 

\textbf{Example:} A computer password consists of 6 symbols from A = {0,..,9} and B = {a,...,z}. How many passwords have a repeated symbol? \\

\noindent Applying the subtraction principle, let \(U\) be the set of all possible passwords, and \(A\) be the set of all passwords with repeated symbols. \\

\begin{center}
  Then \(|A| = |U| - |\bar{A}|\), \\
  There are 36 total symbols, so \(|U| = 36!\), \\
  To find \(|\bar{A}|\), or the number of passwords with no repeated symbols, \\
  The first symbol has 36 possibilities, the second has 35, and so on. So \(|\bar{A}| = \frac{36!}{30!}\), \\
  Therefore, \(|A| = 36! - \frac{36!}{30!}\) \\
\end{center}

\pagebreak
\section*{Permutations of Sets}
Let \(r\) be a positive integer. An \(r\)-permutation of a set \(S\) of \(n\) elements, we get an ordered arrangement of \(r\) of the \(n\) elements. For example, if \(S\) = \{a, b, c\}, then 

\begin{center}
  The three 1-permutations of S are: \\
  \begin{tabular}{c c c}
    a & b & c \\
  \end{tabular}

  The six 2-permutations of S are: \\
  \begin{tabular}{c c c c c c}
    ab & ac & bc & ba & ca & cb \\
  \end{tabular}

  The six 3-permutations of S are: \\
  \begin{tabular}{c c c c c c}
    abc & acb & bac & bca & cab & cba \\
  \end{tabular}
\end{center} 

\noindent The number of \(r\)-permutation of an \(n\)-element set is denoted by \(P(n, r)\).

\subsection*{Theorem 1}

For positive integers \(r\) and \(n\) where \(r \leq n\),
\begin{center}
  \(P(n, r) = n * (n-1) *(n-2) * \dots * (n-r+1) = \frac{n!}{(n-r)!}\) \\
\end{center}

\noindent This is used to find the number of ways to arrange \(r\) objects from a set of \(n\) objects without repetition. \\

\end{document}